From Precipitation to Prediction: Integrating ConvLSTM Models for Comprehensive River Runoff Forecasting

### 2024-04-16 paper/paper.qmd
More text in results

For the evaluation of the model performance we consider the period 1979 to 2011. For this period no bias correction was applied to the orignal E-HYPE dataset. We chose a split of 80% training data, 10% validation data to evaluate the performance of the model during training, and 10% training data that is finally used to evaluate the performance of the model after training.

The accuracy of the model is further displayed in @fig-statistical-evaluationNN . The correlation is close to 1 and the residuals show the error of the model is below 1% of the original data.

![](images/paste-13.png)

@fig-PerformanceNeuralNetworkRunoff shows the performance of the model using the test dataset. The predicted total river runoff for the Baltic Sea is closely matching the original data. Zooming in on the largest individual rivers (lower panels) it can be seen that that also the prediction of the inidivual rivers is close to the original data.

@@

Lastly, we evaluated the performance of the runoff model by incorporating the predicted river runoff as forcing into the ocean model MOM5. This provides a robust validation of the runoff model against more complex real world conditions. This allows us to ensure that the predictions accurately reflect the impact of the river discharge on the ocean dynamics, validating the temporal and spatial variability of the the river discharge. @fig-by15 shows the salinity comparison between the original E-HYPE river runoff and the predicted river runoff at BY15 - a central stations in the Baltic Sea.

>>>>

The model was trained with daily data for the period 1979 to 2011, as this period represents the mapping between the original atmospheric forcing and E-HYPE without further bias correction to match observations. The data was then divided into randomly chosen splits of 80% training data, 10% validation data to evaluate the performance of the model during training, and 10% training data that is finally used to evaluate the performance of the model after training. The model was trained for 400 epochs and the model weights with the lowest mean squared error during training have been stored.

The accuracy of the model is displayed in Figure @fig-statistical-evaluationNN. For this evaluation, the test dataset was utilized. The left panel presents a scatter plot comparing the original E-HYPE data with the predicted river runoff, revealing a high correlation of 0.997. The middle panel illustrates the relative prediction error in relation to the original E-HYPE data, indicating that, on daily timescales, the model can predict river runoff with an accuracy of $\pm$ 5%. Partly, this may be attributed to the use of a different atmospheric dataset than that what was used to force the original E-HYPE model. However, by applying a rolling-mean with a 5-day window, the prediction error is reduced to less than 1%. Lastly, the right panel demonstrates that the distribution of residuals follows a Gaussian shape, suggesting that our model does not exhibit bias by systematically over or underestimating river runoff values.

![](../src/figures/paper_error_metrics.png)

@fig-PerformanceNeuralNetworkRunoff shows the predicted and the original river runoff using the test dataset. The predicted total river runoff for the Baltic Sea is closely matching the original data. Zooming in on the largest individual rivers (lower panels) it can be seen that that also the prediction of the individual rivers is close to the original data.

@@

Lastly, we evaluated the performance of the runoff model by incorporating the predicted river runoff as runoff forcing into the ocean model MOM5. This provides a robust validation of the runoff model against more complex real world conditions. This allows us to ensure that the predictions accurately reflect the impact of the river discharge on the ocean dynamics, validating the temporal and spatial variability of the the river discharge. @fig-by15 shows the salinity comparison between the original E-HYPE river runoff and the predicted river runoff at BY15 - a central stations in the Baltic Sea. It can be seen that the model simulation using the predicted river runoff by the ConvLSTM is closely mirroring the control simulation. The left panel shows the surface salinity, representing the high-frequency variations in the salinity, which is heavily affected by river runoff. The right panel shows the bottom salinity which can be viewed as a low-pass filter in the Baltic Sea, which is also closely mirrored by the ConvLSTM predictions.
